---
title: "WaDE Mapping Overview"
author: "Matthew J Berens"
date: "2023-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/GitHub/WaDE-Theme-1")
```

```{r, echo = FALSE, include=FALSE}
source("code/0-packages.R")
```

### Introduction and Data Access
The boundary of the EFPC watershed is accessed by the `get_huc()` function in the `nhdplusTools` package. The remaining shapefiles were accessed externally through the NHDPlus dataset and manually processed in QGIS. At this point, NHDPlus data for the EFPC watershed are not available through the `nhdplusTools` package.
```{r, echo = FALSE, include=TRUE, message=FALSE, warning = FALSE}
EFPC_HUC <- st_set_crs(get_huc(id = "060102070302", buffer = 0.5, type = "huc12"), 4326)
EFPC_catchments <- st_set_crs(read_sf("GIS_data/EFPC Catchments.shp"), 4326)
EFPC_fl <- st_set_crs(read_sf("GIS_data/EFPC Flowlines.shp"), 4326)
EFPC_main <- st_set_crs(read_sf("GIS_data/East Fork Poplar Creek.shp"), 4326)
```

Land use data can be directly obtained from the National Land Classification Database (NLCD) through the `get_nlcd()` function in the `FedData` package. The data are available for select years from 2001-2019. The year can be selected by changing the `year = ####` option in the `get_nlcd()` function.
```{r, include = TRUE, echo = FALSE, message = FALSE}
nlcd_2019 <- FedData::get_nlcd(template = EFPC_HUC, label = "EFPC NLCD 2019", year = 2019)
```


```{r, echo = FALSE, message=FALSE, include=FALSE, warning=FALSE}
#This block assigns the correct color scheme to the data based on the NLCD color legend.
nlcd_2019 <- raster::as.factor(nlcd_2019)
nlcd_2019@data@attributes[[1]] <- as.data.frame(nlcd_colors())
data <- nlcd_2019@data@attributes[[1]] %>%
  dplyr::left_join(nlcd_colors()) %$%
  Color
```

The WaDE SFA sampling locations are stored in a csv in the GitHub directory. The points must be imported and converted to a spatial object with the `st_as_sf()` function. Ensure that the CRS is set to 4326.
```{r}
WaDE_sites <- read.csv("raw/WaDE SYNOPTIC_locations.csv") %>%
   st_as_sf(coords = c('longitude', 'latitude')) %>%
    st_set_crs(4326)
```


### Overview of Study Area
The map below visualizes the study area for the WaDE SFA. The sampling locations are colorized based on their major tributary catchment. The EFK points have been separated into a unique class. The NLCD data can be toggled on/off from the legend.
```{r, message=FALSE, warning=FALSE, include = TRUE, echo=FALSE}
mapview(WaDE_sites, zcol = "network", layer.name = "Network", alpha.region = 1, label = "site_name", col.regions = RColorBrewer::brewer.pal(length(unique(WaDE_sites$network)), "Spectral")) +
mapview(EFPC_catchments, alpha.regions = 0.2, col.region = "lightgreen", color = "black", legend = FALSE, lwd=1) +
mapview(EFPC_fl, color = "blue", legend = FALSE, lwd = 1) +
mapview(EFPC_main, color = "blue", legend = FALSE, lwd = 2) +
mapview(nlcd_2019, col.regions = data, att = "Class", layer.name = "Landcover Class", label = FALSE, legend = TRUE, alpha = 0.5, hide = TRUE)

```


Now, we can determine which catchment each of the WaDE are in and update the `WaDE_sites` data frame to include the catchment data
```{r, message = FALSE, warning = FALSE}
WaDE_sites <- st_intersection(WaDE_sites, EFPC_catchments)
```


### Exploring Land Use in the EFPC Watershed
```{r, echo = TRUE}
nlcd_colors()
```


```{r, include = FALSE, echo = FALSE, message = FALSE, warning=FALSE}

#This code calculates the zonal statistics for each of the EFPC catchments. The results are exported as a df that is saved as 

landuse_fracs <- exact_extract(nlcd_2019, EFPC_catchments, fun = 'frac', force_df = TRUE)

#Assign the catchments their correct OBJECTID
landuse_fracs$OBJECTID <- EFPC_catchments$OBJECTID

merged <- merge(EFPC_catchments, landuse_fracs, by = "OBJECTID") %>%
  dplyr::select(-c(NHDPlusID, SourceFC, GridCode, VPUID, SHAPE_Leng, SHAPE_Area)) %>%
  mutate(Developed = frac_21 + frac_22 + frac_23 + frac_24,
         Water = frac_11,
         Forest = frac_41 + frac_42 + frac_43,
         Shrub = frac_52,
         Grassland = frac_71,
         Pasture = frac_81,
         Wetland = frac_90 + frac_95,
         Herbaceous = frac_52 + frac_71 + frac_81)
 
sites <- st_intersection(merged, WaDE_sites)
```

```{r, include = TRUE, echo = FALSE, message = FALSE, warning=FALSE}
mapview(sites, layer.name = "Sampling Locations", zcol = "network", alpha = 1, label = "site_name", col.regions = RColorBrewer::brewer.pal(length(unique(sites$network)), "Spectral"))+
mapview(EFPC_fl, color = "blue", legend = FALSE, lwd = 1) +
mapview(EFPC_main, color = "blue", legend = FALSE, lwd = 2) +

mapview(merged, zcol = "Developed", layer.name = "Developed", col.regions = RColorBrewer::brewer.pal(10, "Reds"), hide = TRUE) +
mapview(merged, zcol = "Forest", layer.name = "Forested", col.regions = RColorBrewer::brewer.pal(10, "Greens"), hide = TRUE) +
mapview(merged, zcol = "Wetland", layer.name = "Wetlands", col.regions = RColorBrewer::brewer.pal(10, "Blues"), hide = TRUE) +
mapview(merged, zcol = "Herbaceous", layer.name = "Herbaceous", col.regions = RColorBrewer::brewer.pal(10, "Oranges"), hide = TRUE)

```

### Synoptic Survey Sampling Results
The sampling data are shown below.
```{r, include = FALSE, echo = FALSE, message=F,warning=F}

Apr2023_synoptic <- read.csv("processed/SYNOPTIC_2023-04-12.csv")
Apr2023_synoptic <- merge(Apr2023_synoptic, sites) %>%
  st_as_sf()

June2023_synoptic <- read.csv("processed/SYNOPTIC_2023-06-23.csv")
June2023_synoptic <- merge(June2023_synoptic, sites)


```


```{r April 2023 Map, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}


map_Apr2023 <-
  mapview(subset(Apr2023_synoptic, parameter == "depth"), layer.name = "Flow Status", zcol = "flow_status", alpha.region = 1, label = "site_name", col.regions = c("red", "#57cc99","#ffd60a" )) +
  #mapview(subset(Apr2023_synoptic, parameter == "SO4"), layer.name = "Sulfate", zcol = "result_value", alpha.region = 1, label = "site_name", col.regions = RColorBrewer::brewer.pal(5, "Reds"), hide = TRUE) +
  mapview(subset(Apr2023_synoptic, parameter == "DOC"), layer.name = "DOC", zcol = "result_value", alpha.region = 1, label = "site_name", col.regions = RColorBrewer::brewer.pal(5, "Greys"), hide = TRUE) +
  mapview(EFPC_fl, color = "blue", legend = FALSE, lwd = 1) +
  mapview(EFPC_main, color = "blue", legend = FALSE, lwd = 2) +
  mapview(EFPC_HUC, alpha.regions = 0, col.region = "lightgreen", color = "darkgreen", legend = FALSE, lwd=2) 

```


```{r June 2023 Map, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}


map_June2023 <-
  mapview(subset(June2023_synoptic, parameter == "depth"), layer.name = "Flow Status", zcol = "flow_status", alpha.region = 1, label = "site_name", col.regions = c("red", "#57cc99","#ffd60a" )) +
  #mapview(subset(June2023_synoptic, parameter == "SO4"), layer.name = "Sulfate", zcol = "result_value", alpha.region = 1, label = "site_name", col.regions = RColorBrewer::brewer.pal(5, "Reds"), hide = TRUE) +
  mapview(subset(June2023_synoptic, parameter == "DOC"), layer.name = "DOC", zcol = "result_value", alpha.region = 1, label = "site_name", col.regions = RColorBrewer::brewer.pal(5, "Greys"), hide = TRUE) +
  mapview(EFPC_fl, color = "blue", legend = FALSE, lwd = 1) +
  mapview(EFPC_main, color = "blue", legend = FALSE, lwd = 2) +
 mapview(EFPC_HUC, alpha.regions = 0, col.region = "lightgreen", color = "darkgreen", legend = FALSE, lwd=2) 

map_Apr2023 | map_June2023
```



```{r, message=FALSE, warning = FALSE, fig.cap="Stuff"}

June2023_synoptic %>%
  subset(parameter == "DIC") %>%
  ggplot(aes(y = result_value)) +
  geom_point(aes(x = frac_21), shape = 21, fill = "#7678ed") +
  geom_smooth(aes(x = frac_21), method = lm, se = FALSE, color = "#7678ed") +
  geom_point(aes(x = frac_22), shape = 21, fill = "#f7b801") +
  geom_smooth(aes(x = frac_22), method = lm, se = FALSE, color = "#f7b801") +
  geom_point(aes(x = frac_23), shape = 21, fill = "#f18701") +
  geom_smooth(aes(x = frac_23), method = lm, se = FALSE, color = "#f18701") +
  geom_point(aes(x = frac_24), shape = 21, fill = "#f35b04") +
  geom_smooth(aes(x = frac_24), method = lm, se = FALSE, color = "#f35b04") +
  geom_smooth(aes(x = Developed), method = lm, se = FALSE, color = "black") +
  labs(x = "Developed Land Cover (%)", y = "DIC (mg/L)") +
  #FFD0D0, #FF8B8B, #FF4545, #FF0000

  theme_mb1()

```