---
title: "EFPC Geospatial"
author: "Matthew Berens"
date: "2023-11-30"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/GitHub/WaDE-Theme-1")
setwd("~/Documents/GitHub/WaDE-Theme-1")
```

```{r, echo = FALSE, include=FALSE}
source("code/0-packages.R")
```

### Introduction and Data Access
The boundary of the EFPC watershed is accessed by the `get_huc()` function in the `nhdplusTools` package. The remaining shapefiles were accessed externally through the NHDPlus dataset and manually processed in QGIS. At this point, NHDPlus data for the EFPC watershed are not available through the `nhdplusTools` package.
```{r, echo = TRUE, include=TRUE, message=FALSE, warning = FALSE}
EFPC_HUC <- st_set_crs(get_huc(id = "060102070302", buffer = 0.5, type = "huc12"), 4326)
EFPC_catchments <- st_set_crs(read_sf("GIS_data/EFPC Catchments.shp"), 4326)
EFPC_fl <- st_set_crs(read_sf("GIS_data/EFPC Flowlines.shp"), 4326)
EFPC_main <- st_set_crs(read_sf("GIS_data/East Fork Poplar Creek.shp"), 4326)
```

Land use data can be directly obtained from the National Land Classification Database (NLCD) through the `get_nlcd()` function in the `FedData` package. The data are available for select years from 2001-2019. The year can be selected by changing the `year = ####` option in the `get_nlcd()` function.
```{r, include = TRUE, echo = TRUE, message = FALSE}
nlcd_2019 <- FedData::get_nlcd(template = EFPC_HUC, label = "EFK NLCD 2019", 
                               year = 2019)
```


This chunk generates the correct color scheme for the NLCD data.
```{r, echo = TRUE, message=TRUE, include=FALSE, warning=FALSE}
#This block assigns the correct color scheme to the data based on the NLCD color legend.
nlcd_2019 <- raster::as.factor(nlcd_2019)
nlcd_2019@data@attributes[[1]] <- as.data.frame(nlcd_colors())
data <- nlcd_2019@data@attributes[[1]] %>%
  dplyr::left_join(nlcd_colors()) %$%
  Color
```

The WaDE SFA sampling locations are stored in a csv in the GitHub directory. The points must be imported and converted to a spatial object with the `st_as_sf()` function. Ensure that the CRS is set to 4326.
```{r}
WaDE_sites <- read.csv("raw/WaDE SYNOPTIC_SITES.csv") %>%
   st_as_sf(coords = c('longitude', 'latitude')) %>%
    st_set_crs(4326)
```


### Overview of Study Area
Here is a simple map  that uses the `mapview` function to visualize the study area for the WaDE SFA. I colorized the sampling locations based on their tributary subcatchment. The NLCD data can be toggled on/off from the legend. Each of the shapefiles for the catchments, flowlines, etc. are added as separate mapview objects.
```{r, message=TRUE, warning=FALSE, include = TRUE, echo=TRUE}
mapview(WaDE_sites, zcol = "network", layer.name = "Sampling Locations", alpha.region = 1, label = "site_name", col.regions =   RColorBrewer::brewer.pal(length(unique(WaDE_sites$network)), "Spectral")) +
mapview(EFPC_catchments, alpha.regions = 0.2, col.region = "lightgreen", color = "black", legend = FALSE, lwd=1, layer.name = "EFK Catchments") +
mapview(EFPC_fl, color = "blue", legend = FALSE, lwd = 1, layer.name = "NHD Flowlines") +
mapview(EFPC_main, color = "blue", legend = FALSE, lwd = 2, layer.name = "EFK Main Branch") +
mapview(nlcd_2019, col.regions = data, att = "Class", layer.name = "Landcover Class", label = FALSE, legend = TRUE, alpha = 0.5, hide = TRUE)

```



This function overlaps the WaDE sampling locations with the catchment boundary shapefile to determine which catchment each sampling location is in. The `WaDE_sites` data frame is then updataed to include the catchment data
```{r, message = TRUE, warning = FALSE}
WaDE_sites <- st_intersection(WaDE_sites, EFPC_catchments)
```


### Land Cover Calculations
This chunk calculates the catchment-specific land use percentages from the NLCD and then bins them into larger categorical groupings (e.g., developed, water, forested). I save them as geopackages so they can be easily accessed later. Sometimes saving things as .shp files in R messes up the names of the data frames.
```{r, include = TRUE, echo = TRUE, message = FALSE, warning=FALSE, results = 'hide'}

#This code calculates the zonal statistics for each of the EFPC catchments. 
#The results are exported as a df that is saved as landuse_fracs

landuse_fracs <- exact_extract(nlcd_2019, EFPC_catchments, fun = 'frac', 
                               force_df = TRUE)

#Assign the catchments their correct OBJECTID
landuse_fracs$OBJECTID <- EFPC_catchments$OBJECTID

EFPC_catchments_landuse <- merge(EFPC_catchments, landuse_fracs,
                                 by = "OBJECTID") %>%
                                dplyr::select(-c(NHDPlusID, SourceFC, 
                                GridCode, VPUID, SHAPE_Leng, SHAPE_Area)) %>%
  mutate(Developed = frac_21 + frac_22 + frac_23 + frac_24,
         Water = frac_11,
         Forest = frac_41 + frac_42 + frac_43,
         Shrub = frac_52,
         Grassland = frac_71,
         Pasture = frac_81,
         Wetland = frac_90 + frac_95,
         Herbaceous = frac_52 + frac_71 + frac_81,
         OBJECTID = as.character(OBJECTID))
 

#Merge the catchment land use percentages with the sampling site locations
WaDE_sites_landuse <- st_intersection(EFPC_catchments_landuse, WaDE_sites) %>%
  dplyr::select(-c(AreaSqKm.1))

#Export combined land use calculations with sampling locations
st_write(EFPC_catchments_landuse, "GIS_data/EFPC_catchments_landuse.gpkg", 
         append = FALSE)

st_write(WaDE_sites_landuse, "GIS_data/WaDE_sites_landuse.gpkg", append = FALSE)
```



Next, I manually grouped the subcatchments together into one larger catchment that encompasses and entire upstream portion of the tributary network for each site
```{r, message = FALSE, warning = FALSE, results = 'hide'}
catchment_bins <- rbind(
data.frame("site_name" = "NBO1", "OBJECTID" = "89458"),
data.frame("site_name" = "NBO2", "OBJECTID" = c("238691","100959")),
data.frame("site_name" = "NBO3", "OBJECTID" = "100959"),
data.frame("site_name" = "NBO4", "OBJECTID" = "102476"),
data.frame("site_name" = "NBO5", "OBJECTID" = c("231934","242844", "98890","99751","99669")),
data.frame("site_name" = "NBO6", "OBJECTID" = "101341"),
data.frame("site_name" = "MTN4", "OBJECTID" = "90316"),
data.frame("site_name" = "MTN3", "OBJECTID" = c("90316","257888","90620")),
data.frame("site_name" = "MTN2", "OBJECTID" = c("90316","257888","90620")),
data.frame("site_name" = "MTN1.3", "OBJECTID" = "93594"),
data.frame("site_name" = "MTN1.2", "OBJECTID" = c("90316","257888","90620")),
data.frame("site_name" = "MTN1.1", "OBJECTID" = c("245543","93594","90316","257888","90620")),
data.frame("site_name" = "MCA4", "OBJECTID" = c("238228","91132","90546")),
data.frame("site_name" = "MCA3.3", "OBJECTID" = c("89979")),
data.frame("site_name" = "MCA3.2", "OBJECTID" = c("91501")),
data.frame("site_name" = "MCA3.1", "OBJECTID" = c("231863", "91501", "89979")), 
data.frame("site_name" = "MCA2" , "OBJECTID" = c("238322","238228","91132","90546","231863", "91501","89979")),
data.frame("site_name" = "MCA1" , "OBJECTID" = c("238322","238228","91132","90546","231863", "91501", "89979","90467", "90031")),
data.frame("site_name" = "WBK1" , "OBJECTID" = c("245547","94989")),
data.frame("site_name" = "WBK2" , "OBJECTID" = c("245547","94989")),
data.frame("site_name" = "WBK3" , "OBJECTID" = c("93685")),
data.frame("site_name" = "BSL8" , "OBJECTID" = c("81824")),
data.frame("site_name" = "BSL7" , "OBJECTID" = c("73872")),
data.frame("site_name" = "BSL6" , "OBJECTID" = c("85970")),
data.frame("site_name" = "BSL5.3" , "OBJECTID" = c("258680")),
data.frame("site_name" = "BSL5.2" , "OBJECTID" = c("72273")),
data.frame("site_name" = "BSL5.1" , "OBJECTID" = c("238341","258680","72273")),
data.frame("site_name" = "BSL4.3" , "OBJECTID" = c("73872")),
data.frame("site_name" = "BSL4.2" , "OBJECTID" = c("81824")),
data.frame("site_name" = "BSL4.1" , "OBJECTID" = c("245291","81824","73872")),
data.frame("site_name" = "BSL3.3" , "OBJECTID" = c("238341","258680","72273")),
data.frame("site_name" = "BSL3.2" , "OBJECTID" = c("85970")),
data.frame("site_name" = "BSL3.1" , "OBJECTID" = c("85970","238341","258680","72273")),
data.frame("site_name" = "BSL2.3" , "OBJECTID" = c("85970","238341","258680","72273")),
data.frame("site_name" = "BSL2.2" , "OBJECTID" = c("245291","81824","73872")),
data.frame("site_name" = "BSL2.1" , "OBJECTID" = c("85970","238341","258680","72273","245291","81824","73872")),
data.frame("site_name" = "BSL1.3" , "OBJECTID" = c("85970","238341","258680","72273","245291","81824","73872")),
data.frame("site_name" = "BSL1.2" , "OBJECTID" = c("241710")),
data.frame("site_name" = "BSL1.1" , "OBJECTID" = c("85970","238341","258680","72273","245291","81824","73872","95129","241710"))
)

catchments_binned <- merge(EFPC_catchments, catchment_bins, by = "OBJECTID") %>% 
    group_by(site_name) %>%
    summarise(geometry = sf::st_union(geometry)) %>%
    ungroup() %>%
    st_as_sf() %>%
    mutate(AreaSqKm = st_area(geometry)/1000000)

#This code calculates the zonal statistics for each of the EFPC catchments. 
#The results are exported as a df that is saved as landuse_bins
landuse_bins <- exact_extract(nlcd_2019, catchments_binned, fun = 'frac', 
                              force_df = TRUE)

landuse_bins$site_name = catchments_binned$site_name

EFPC_catchments_binned <- merge(catchments_binned, landuse_bins, 
                                by = "site_name") %>%
  
  mutate(Developed = frac_21 + frac_22 + frac_23 + frac_24,
         Water = frac_11,
         Forest = frac_41 + frac_42 + frac_43,
         Shrub = frac_52,
         Grassland = frac_71,
         Pasture = frac_81,
         Wetland = frac_90 + frac_95,
         Herbaceous = frac_52 + frac_71 + frac_81)

#Export combined land use calculations with sampling locations
st_write(EFPC_catchments_binned, "GIS_data/EFPC_LU_catchments_binned.gpkg",
         append = FALSE)

```




