---
title: "WaDE June Synoptic Survey"
author: "Matthew J Berens"
date: "2023-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/GitHub/WaDE-Theme-1")
setwd("~/Documents/GitHub/WaDE-Theme-1")
```

```{r, echo = FALSE, include=FALSE}
source("code/0-packages.R")
```

### Introduction and Data Access
The boundary of the EFPC watershed is accessed by the `get_huc()` function in the `nhdplusTools` package. The remaining shapefiles were accessed externally through the NHDPlus dataset and manually processed in QGIS. At this point, NHDPlus data for the EFPC watershed are not available through the `nhdplusTools` package.
```{r, echo = FALSE, include=TRUE, message=FALSE, warning = FALSE}
EFPC_HUC <- st_set_crs(get_huc(id = "060102070302", buffer = 0.5, type = "huc12"), 4326)
EFPC_catchments <- st_set_crs(read_sf("GIS_data/EFPC Catchments.shp"), 4326)
EFPC_fl <- st_set_crs(read_sf("GIS_data/EFPC Flowlines.shp"), 4326)
EFPC_main <- st_set_crs(read_sf("GIS_data/East Fork Poplar Creek.shp"), 4326)
EFPC_catchments <- st_set_crs(read_sf("GIS_data/EFPC Catchments.shp"), 4326)
WaDE_sites_landuse <- st_set_crs(read_sf("GIS_data/WaDE_sites_landuse.gpkg"), 4326)
EFPC_LU_catchments_binned <- st_set_crs(read_sf("GIS_data/EFPC_LU_catchments_binned.gpkg"), 4326)
```


### Overview of Study Area
The map below visualizes the study area for the WaDE SFA. The sampling locations are colorized based on their major tributary catchment.
```{r, message=FALSE, warning=FALSE, include = TRUE, echo=FALSE}
mapview(WaDE_sites_landuse, zcol = "network", layer.name = "Sampling Locations", alpha.region = 1, label = "site_name", col.regions = RColorBrewer::brewer.pal(length(unique(WaDE_sites_landuse$network)), "Spectral")) +
mapview(EFPC_catchments, alpha.regions = 0.2, col.region = "lightgreen", color = "black", legend = FALSE, lwd=1, layer.name = "EFK Catchments") +
mapview(EFPC_fl, color = "blue", legend = FALSE, lwd = 1, layer.name = "NHD Flowlines") +
mapview(EFPC_main, color = "blue", legend = FALSE, lwd = 2, layer.name = "EFK Main Branch")

```


### Synoptic Survey Sampling Results
The sampling data are shown below.
```{r, include = FALSE, echo = FALSE, message=F,warning=F}
Apr2023_synoptic <- read.csv("processed/SYNOPTIC_2023-04-12.csv")
Apr2023_synoptic <- merge(Apr2023_synoptic, WaDE_sites_landuse) %>%
  st_as_sf()

June2023_synoptic <- read.csv("processed/SYNOPTIC_2023-06-23.csv")
June2023_synoptic <- merge(June2023_synoptic, WaDE_sites_landuse) %>%
  st_as_sf()
```


```{r April 2023 Map, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}
map_Apr2023 <-
  mapview(EFPC_fl, color = "blue", legend = FALSE, lwd = 1, layer.name = "NHD Flowlines") +
  mapview(EFPC_main, color = "blue", legend = FALSE, lwd = 2, layer.name = "EFK Main Branch") +
  mapview(EFPC_HUC, alpha.regions = 0, col.region = "lightgreen", color = "darkgreen", legend = FALSE, lwd=2) +
  mapview(subset(Apr2023_synoptic, parameter == "depth"), layer.name = "Flow Status", zcol = "flow_status", alpha.region = 1, label = "site_name", col.regions = c("red", "#57cc99","#ffd60a" )) 
```


```{r June 2023 Map, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}
map_June2023 <-
  mapview(EFPC_fl, color = "blue", legend = FALSE, lwd = 1, layer.name = "NHD Flowlines") +
  mapview(EFPC_main, color = "blue", legend = FALSE, lwd = 2, layer.name = "EFK Main Branch") +
  mapview(EFPC_HUC, alpha.regions = 0, col.region = "lightgreen", color = "darkgreen", legend = FALSE, lwd=2) +
  mapview(subset(June2023_synoptic, parameter == "depth"), layer.name = "Flow Status", zcol = "flow_status", alpha.region = 1, label = "site_name", col.regions = c("red", "#57cc99","#ffd60a")) 
  

map_Apr2023 | map_June2023
```


```{r June 2023 Map, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}
  mapview(EFPC_fl, color = "blue", legend = FALSE, lwd = 1, layer.name = "NHD Flowlines") +
  mapview(EFPC_main, color = "blue", legend = FALSE, lwd = 2, layer.name = "EFK Main Branch") +
  mapview(EFPC_HUC, alpha.regions = 0, col.region = "lightgreen", color = "darkgreen", legend = FALSE, lwd=2) +
  mapview(subset(June2023_synoptic, parameter == "Ca"), layer.name = "Silicon", zcol = "result_value", alpha.region = 1, label = "site_name", col.regions = RColorBrewer::brewer.pal(20, "Blues")) 
```



```{r, message=FALSE, warning = FALSE, fig.cap="Stuff"}
June2023_synoptic_NObins <- merge(read.csv("processed/SYNOPTIC_2023-06-23.csv"), st_drop_geometry(WaDE_sites_landuse))
June2023_synoptic_bins <- merge(read.csv("processed/SYNOPTIC_2023-06-23.csv"), st_drop_geometry(EFPC_LU_catchments_binned))

Apr2023_synoptic_NObins <- merge(read.csv("processed/SYNOPTIC_2023-04-12.csv"), st_drop_geometry(WaDE_sites_landuse))
Apr2023_synoptic_bins <- merge(read.csv("processed/SYNOPTIC_2023-04-12.csv"), st_drop_geometry(EFPC_LU_catchments_binned))

NObins <- rbind(Apr2023_synoptic_NObins, June2023_synoptic_NObins) %>%
  mutate(binned = "N") %>%
  dplyr::select(-c(NHDPlusID, SourceFC, GridCode, VPUID, SHAPE_Leng, SHAPE_Area, project_name, network, theme, OBJECTID.1, OBJECTID))

bins <- rbind(Apr2023_synoptic_bins, June2023_synoptic_bins) %>%
  mutate(binned = "Y")



#rbind(NObins, bins) %>%
bins %>%
  subset(parameter %in% c("Mg", "Ca", "SpC", "DIC", "DOC", "Cl")) %>%
  #subset(parameter == "SpC") %>%
  ggplot(aes(y = result_value)) +
  geom_smooth(aes(x = Developed*100, color = sample_date), method = lm, se = TRUE) +
  geom_point(aes(x = Developed*100, fill = sample_date), shape = 21, size = 4) +
  labs(x = "Developed Land Cover (%)", y = "Cond. (uS/cm))") +
  theme_mb1() +
  facet_wrap(~parameter, scales = "free_y", nrow = 2) +
  theme(aspect.ratio = 1, legend.position = "none", axis.title = element_text(size = 18), axis.text = element_text(size = 18))


ggsave(plot = last_plot(), "output/comparisons.png", width = 12, height = 9, units = "in")
  
```